### 2015-09-28 GR_param.tex

\section{ Gutenberg-Richter Parameters}
For each of the three seismic regions, Gutenberg-Richter parameters and Max magnitude ($M_{max}$) were calculated using Seismic Hazard Assessment code (HA3) \citep{kijko2004}. The regional maximum magnitude for each region is estimated by the \citet{Kijko1989} method, which is implemented in the HA3 package. For smoothed seismicity areas, $b-value$ is assumed constant. The $a-value$ can vary spatially and is determined by counting earthquakes above $M$ 3.0 in grid cells.\\
\noindent
\citet{Karimiparidari2013} applied the Maximum Curvature (MAXC) technique \citep{Wyss1999, Wiemer2000} by ZMap \citep{Wiemer2001} to calculate the level of completeness of instrumental part of the catalog. In this study we use those magnitudes of completeness as the magnitude threshold in the calculation of the seismicity parameters. We also used the seismicity parameters of this study \citep{Karimiparidari2013} as the priory information in HA3 code. The updated values are displayed in Table 2. \\
\noindent
Following the \citet{Karimiparidari2013}, we assume the catalog is complete for earthquakes with magnitude 4.5, 4.4, and 4.5 for Azerbaijan, Alborz and Kopeh-Dagh tectonic seismic regions, respectively. The completeness of each region can be seen easily from the scatter plot of completeness test.

\begin{table}[h]
\centering
\caption{Seismicity parameters for three tectonic seismic regions.}
    \begin{tabular}{ccccc}
    ~                   & $b-value$            & $\lambda$                  & $M_{max}$ Calculated & $M_{max}$ obseved \\ \hline
    Azerbaijan    & 1.06 $\pm$ 0.03  & (1.827  $\pm$ 0.216) & 7.78  $\pm$ 0.26           & 7.7          \\ \hline
    Alborz           & 1.1  $\pm$ 0.03   & (2.290  $\pm$ 0.282) & 7.91  $\pm$ 0.27           & 7.7          \\ \hline
    Kopeh Dagh & 0.97  $\pm$ 0.04 & (1.716  $\pm$ 0.243) & 7.67  $\pm$ 0.26           & 7.6          \\
    \end{tabular}

\end{table}