### 2015-09-28 completness_test.tex

\section{Completeness tests}
For the smoothed seismicity method, completeness of each magnitude in the catalog is an important factor. To access catalog completeness, according to \citet{Frankel1995}, we made plots of the cumulative number of events against time for different regions.   Fig.~\ref{fig:comptest} shows the completeness of data for each magnitude threshold in three different regions.
A uniform rise of cumulative number of earthquake in each magnitude big, defines the threshold for catalog completeness.

\begin{figure} [H]
\centering
\includegraphics[scale=0.3]{figures/pdf/Figure4.pdf} 
\caption{Completeness test for earthquakes occurred in the three study regions. Red lines represent $M_w > 7$ }
\label{fig:comptest}
\end{figure}

We assume the catalog for earthquakes with magnitude greater than $M_w=7$ is complete from the first historical earthquake report. Table 1 presents the completeness of catalog of each region for different moment magnitudes.

\begin{table}[h]
\centering
\caption{Completeness threshold of three tectonic seismic regions.}
\begin{tabular}{cccc}
 ~           & Azerbaijan & Alborz & Kopeh Dagh \\ \hline
3-4         & 2004       & 2004   & 2005       \\ \hline
4-5         & 1960       & 1960   & 1960       \\ \hline
5-6         & 1900       & 1900   & 1900       \\ \hline
6-7         & 1844       & 1809   & 1810         \\ \hline
7 $< $    &  85          & -400    & 10         \\ \hline
\end{tabular}
\end{table}