### 2015-09-28 main.tex

\documentclass{article}

\usepackage{textcomp}
\usepackage{amsmath}    % need for subequations
\usepackage{graphicx}   % need for figures
\usepackage{verbatim}   % useful for program listings
\usepackage{color}      % use if color is used in text
\usepackage{subfigure}  % use for side-by-side figures
\usepackage{hyperref}   % use for hypertext links, including those to external documents and URLs
\usepackage{cite}
\usepackage{float}
\usepackage{natbib}
\usepackage[utf8]{inputenc}
\usepackage{simplemargins}
\usepackage{setspace}
\usepackage{multirow}
\usepackage{url}

\usepackage{etoolbox}
\usepackage{lipsum}

\setallmargins{1in}
\onehalfspacing

\linespread{1.8} % line space

\patchcmd{\thebibliography}{\section*}{\section*}{}{}

\usepackage{xcolor}
\hypersetup{
    colorlinks,
    linkcolor={red!50!black},
    citecolor={blue!50!black},
    urlcolor={blue!80!black}
}

\title{Seismic Hazard Estimation of Northern Iran Using Smoothed Historical Seismicity}

\begin{document}

\maketitle

\newpage

\begin{center}
\textbf{Naeem Khoshnevis$^*$} \\ Graduate Research Assistant \\  University of Memphis \\ Center for Earthquake Research and Information \\   nkhshnvs@memphis.edu\\   3890 Central Ave. Memphis TN 38152 \\ Phone (901) 678-2007\\
\vspace{10 mm}
 \textbf{Shima Azizzadehroodpish}  \\ Graduate Research Assistant \\ University of Memphis \\ Center for Earthquake Research and Information \\ szzzdhrd@memphis.edu\\  
\vspace{10 mm}
\textbf{Chris H Cramer}   \\ Research  Associate  Professor \\  University of Memphis \\ Center for Earthquake Research and Information \\ ccramer@memphis.edu \\
\vspace{10 mm}
\textbf{Ricardo Taborda}   \\  Assistant  Professor \\  University of Memphis \\ Department of Civil Engineering and \\ Center for Earthquake Research and Information \\ rtbrdros@memphis.edu\\

\end{center}

\newpage

\begin{abstract}
\input{abstract}
\end{abstract}

\smallskip
\noindent \textbf{Keywords: } Seismic hazard - Spatially smoothed seismicity - Northern Iran

\input{content}

\newpage

{
\renewcommand{\bibname}{}
\bibliographystyle{ascelike}
\bibliography{PSHA}
}

\end{document}