### 2016-01-11 springer_main.tex

\RequirePackage{fix-cm}

%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
%\documentclass[natbib,smallextended]{svjour3}       % onecolumn (second format)

\documentclass[natbib,twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{graphicx}
\usepackage[a4paper]{geometry}
\usepackage{xcolor}
%\usepackage[margin=1in]{geometry}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
 \usepackage{aps-bibstyle}  % use this style if you upload to .tex file only a part of Bibtex created bbl.
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
 \journalname{Journal of Seismology}
%
% Definitions for the journal names
\newcommand{\aap}{{Astron. Astrophys.}}
\newcommand{\apj}{{Astrophys. J.}}
\newcommand{\grl}{{Geophys. Res. Lett.}}
\newcommand{\solphys}{{Solar Phys.}}
%
\begin{document}

\title{ Seismic Hazard Estimation of Northern Iran Using Smoothed Historical Seismicity}

\author{Naeem Khoshnevis         \and
             Shima Azizzadehroodpish \and
             Chris H Cramer \and
             Ricardo Taborda}

%\authorrunning{Short form of author list} % if too long for running head

\institute{Naeem Khoshnevis \at
               University of Memphis \\
               Center for Earthquake Research and Information \\
               3890 Central Ave. Memphis TN 38152 \\
              Tel.: +1(901) 678-2007\\
               \email{nkhshnvs@memphis.edu}        
           \and
           Shima Azizzadehroodpish \at
                            University of Memphis \\
               Center for Earthquake Research and Information \\
               Department of Civil Engineering \\
             \email{szzzdhrd@memphis.edu}        
\and
Chris H Cramer \at
                            University of Memphis \\
               Center for Earthquake Research and Information \\
               \email{ccramer@memphis.edu}    
\and
Ricardo Taborda \at
                            University of Memphis \\
               Center for Earthquake Research and Information \\
               \email{rtbrdros@memphis.edu}   
}

\date{Received: date / Accepted: date}

\maketitle

\begin{abstract}
\input{abstract}
\keywords{Seismic hazard \and Spatially smoothed seismicity \and Northern Iran}

\end{abstract}

\input{content}

\newpage
\bibliographystyle{aps-nameyear}      % American Physical Society (APS) style, author-year citations
\bibliography{PSHA}                           % name your BibTeX data base
\nocite{*}

\end{document}