### 2016-01-17 hazard_analysis_results.tex

\section{Hazard Analysis Results}

The PGA hazard maps have been computed for a 10\% and 2\% exceedance probability in 50 years, corresponding to a return period of 475 and 2475 years, from gridded values of historical and instrumental seismic activity. These levels of exceedance are a standard practice in seismic designs \citep{BHRC2014}.\\
\noindent
Even though in this study we are interested in investigating seismic hazard study in the northern Iran, in the smoothed seismicity method we defined the border of the study region broadly enough to ensure that the seismicity outside of the border doesn't affect the study area.  In addition, the border of the influenced area should surround the area of interest as uniformly as possible \citep{Lapajne1997}. Therefore, our study region includes parts of central and eastern Iran and the Zagros tectonic seismic regions.\\
\noindent
We calculated the $a-values$ for each cell and spatially smoothed over a grid of $0.1 \times 0.1$ in latitude and longitude. We assume magnitude increment as $\Delta_M = 0.1$. In this model, events are not assigned to specific faults and are assumed to be potential seismogenic sources, and are spatially gridded to cells. \\
\noindent
Different correlation distances have been used in different studies, which are generally dependent on the accuracy of the location of the recorded earthquakes. \citet{Frankel1995} and \citet{Boyd2008} assumed the correlation distance to be 50 $km$. \citet{Barani2007} used distance of 25 $km$ based on previous studies that suggested the correlation function of the Alps and Apennines in Italy. \citet{Foteva2006} used 10 and 15 km in a different model. Correlation distance is a very sensitive parameter in smoothed seismic hazard maps, and it is a factor of uncertainty in earthquake location. In the Iranian earthquake catalog, the magnitude uncertainties were assumed to be in the interval of $\pm$ 0.25 magnitude units and epicentral errors of $\pm$ 30 km \citep{Zare2012}. Even though these error margins are different for different events in terms of instrumental or historical events, in this study we assumed the correlation distance as 30 km because of the lack of knowledge about other historical events. Regarding the study domain and the 0.1 grid size, the total number of source is grid sites is 11651. Even though \citet{Kalkan2004} derived the attenuation relationship up to 250 km, since \citet{Zafarani2014} used data up to 200 km to evaluate the GMPEs, we set the model to compute the ground motions at distances of less than 200 km.
%Fig.~\ref{fig:10percent} and Fig.~\ref{fig:2percent} show the seismic hazard map based on background seismicity for 10\% and 2\% probability of exceedance in 50 years, respectively. The areas of large probabilistic ground motions clearly coincide with zones with a large number of events with magnitude 3 and larger.

\subsection{Hazard Curve}
We calculated the probability of exceedance of the ground motion from different ground motion levels, separately for each of 5 regions. We add the probability of all regions in order to have the effect of seismicity from all tectonic seismic regions. The results are presented as annual rate of exceedance. Fig.~\ref{fig:hazardcurve} represents the hazard curve of different model for some of important northern cities in Iran.

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.4]{figures/pdf/HazardCurve.pdf} 
\caption{Seismic Hazard Curve of Northern Cities in Iran}
\label{fig:hazardcurve}
\end{figure*}

\subsection{5 Regions}
We pick the peak ground acceleration for 10\% and 2\% probability of exceedance in 50 years which correspond with 0.002 and 0.0004 annual rate of exceedance. 
Fig.~\ref{fig:pga_10_mean} and Fig.~\ref{fig:pga_2_mean} show the results for 10\% and 2\% probability of exceedance in 50 years, respectively. These results are according to the mean value of the attenuation relationship.

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.3]{figures/pdf/pga_10_mean.pdf} 
\caption{Peak ground acceleration for 10\% probability of exceedance in 50 years.}
\label{fig:pga_10_mean}
\end{figure*}

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.3]{figures/pdf/pga_2_mean.pdf} 
\caption{Peak ground acceleration for 2\% probability of exceedance in 50 years.}
\label{fig:pga_2_mean}
\end{figure*}

\subsection{Uniform Region}
Fig.~\ref{fig:pga_10_mean_uniform} and Fig.~\ref{fig:pga_2_mean_uniform} show the results for 10\% and 2\% probability of exceedance in 50 years for the uniform model, respectively.

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.3]{figures/pdf/pga_10_mean_uniform.pdf} 
\caption{Peak ground acceleration for 10\% probability of exceedance in 50 years (uniform model).}
\label{fig:pga_10_mean_uniform}
\end{figure*}

\begin{figure*} [!ht]
\centering
\includegraphics[scale=0.3]{figures/pdf/pga_2_mean_uniform.pdf} 
\caption{Peak ground acceleration for 2\% probability of exceedance in 50 years (uniform model).}
\label{fig:pga_2_mean_uniform}
\end{figure*}

%\subsection{3 Regions}