### 2016-01-31 Conclusion.tex
Conclusion added.

\section{Conclusion}
The result of a new probabilistic seismic hazard assessment for northern Iran using updated seismic catalogs is provided. Seismic hazard is evaluated over entire three regions (Azerbaijan, Alboz, Kopeh-Dagh) as well as part of Central and Zagros tectonic seismic regions using the smoothed background seismicity. In order to determine the seismic hazard in North Iran, different studies have been done. These studies are based on defining faults as a seismic source model. Using faults as a seismic source model could result in accurate values in case the faults are precisely determined. There are considerable uncertainties on fault characteristic, specially on fault length. Instead of using faults as a seismic source, we consider each earthquake event as a point source in the model. We reduce the uncertainty in earthquake location by smoothing the number of earthquake in adjacent cells. We defined models based on considering the whole North Iran as a uniform seismic tectonic regions and 5 different seismic tectonic regions. The differences in catalog completeness, b value, and maximum magnitude of earthquake in these two models resulted in different results. This indicate the idea that dividing regions into subregions is necessary in order to be able to get accurate results. The result are also sensitive to the minimum magnitude. The structures built according to engineering code are supposed to withstand the earthquake with magnitude 5 at the reasonable distance. However, closer $M_w > 4.5$ for those structures and also older masonry structures could cause structural damage. A combination of two models $(M_w > 4.5, M_w>5)$ in each city based on number of engineering and old masonry structures will provide more accurate results. The results are provided for mean peak horizontal acceleration in bedrock. The ground acceleration in soil deposits and inside basin will be bigger than the provided values. This model reduce the epistemic uncertainty regarding the recognizing faults and earthquake location. Combination of this model with a study based on using faults will increase the values for existing non-active seismic fault regions.