### 2016-07-08 main.tex
looks like done

\RequirePackage{fix-cm}

% Document Settings
\documentclass[natbib,twocolumn]{svjour3}

% Flushing
\smartqed

% Packages
\usepackage{graphicx}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{multirow}
\usepackage{marvosym}
\usepackage{textcomp}

% Journal Name
\journalname{Journal of Seismology}

% START: The Document
\begin{document}

% Long Title
\title{Seismic Hazard Estimation of Northern Iran Using Smoothed Seismicity}

% Running Title
\titlerunning{Seismic Hazard of Northern Iran}

% Authors
\author{%
	N.~Khoshnevis \and
	R.~Taborda \and
	S.~Azizzadeh-Roodpish \and
	C.~H.~Cramer
}

% Affiliations
\institute{%
	N.~Khoshnevis \and C.~H.~Cramer \at
	Center for Earthquake Research and Information,
	University of Memphis, 3890 Central Ave., Memphis TN 38152, USA\\
	\and
	R.~Taborda (\Letter) \and S.~Azizzadeh-Roodpish \at
	Department of Civil Engineering, and
	Center for Earthquake Research and Information,
	University of Memphis, 3890 Central Ave., Memphis TN 38152, USA\\
	\email{ricardo.taborda@memphis.edu}  
}

% Date
% \date{Received: date / Accepted: date}

% Title Production
\maketitle

% Abstract
\begin{abstract}
	\input{abstract}
	\keywords{%
		Seismic hazard \and 
		Spatially smoothed seismicity 
		\and Northern Iran
	}
\end{abstract}

% Article Content
\input{content}

% Bibliography
\bibliographystyle{spbasic}
\bibliography{references}
% \nocite{*}

% END: The Document
\end{document}