### 2016-07-19 catalog.tex
implemented last set of changes by cramer

\section{Seismic Catalog}

\input{compilation}

\input{completeness}