### 2016-10-20 introduction.tex

\section{Introduction}

This article presents a new seismic hazard assessment for the northern portion of the Iranian plateau, using a smoothed seismicity approach in combination with an updated seismic catalog and a ground motion prediction equation recently evaluated and found to fit well with observations of peak ground response in the region.

The Iranian plateau is located on the Himalayan-Alpine seismic belt. It is confined between the convergent movements of the Arabian and Eurasian plates, and laterally trapped between the Arabian and eastern Asia-Minor to the west, and the Indian and Eurasian plates to the east \citep{Berberian_1981_Chap}. This entrapment lifts the plateau and gives rise to the three major mountain ranges delimiting Iran's land, the Zagros mountain ranges running from the southeast to the northwest, the Alborz ranges to the north, and the east Iranian ranges \myrevision{(see Fig.~\ref{fig:iran})}. It also defines the seismic activity of the region---one with a long history of large magnitude ($M>7$) earthquakes that are well documented dating back to the \myrevision{eighth} century---which has repeatedly damaged Iran's infrastructure and taken the lives of thousands.

According to this tectonic setting and the geologic provinces of the plateau, Iranian earthquakes have been categorized into various seismic zones, ranging from the definition of four to nine major seismic zones in the more traditional studies \citep[e.g.,][]{Stocklin1968, Takin1972, Berberian1976}, and up to twenty to twenty-three seismotectonic provinces in the most elaborate ones \citep[e.g.,][]{Nowroozi1976, Tavakoli1999}. Contemporary, well-accepted models divide Iran into five major tectonic regions: Azerbaijan-Alborz, Kopeh-Dagh, Zagros, Central-East Iran, and Makran \citep[e.g.,][]{Mirzaei1998}. More recently, in a study aimed to develop a uniform earthquake catalog for Iran, \citet{Karimiparidari2013} further divided the Azerbaijan-Alborz region into two separate seismic zones: Azerbaijan and the Alborz Mountain Range, hereafter referred to simply as Alborz. We adopt the latter definition of six seismic zones and focus our attention on those most relevant to the northern part of the country, Azerbaijan, Alborz and Kopeh-Dagh.

\myrevision{Over the last three decades there have been multiple efforts to estimate the seismic parameters and seismic hazard of Iran, and improve the assessment of seismic risk in the country \citep[e.g.,][]{Tavakoli1999, Moinfar_2000_Chap, Ghodrati2003, Moinfar_2012_WCEE, Khodaverdian_2016_BSSA}. Altogether, these previous studies have helped improve the Iranian Code of Practice for Seismic Resistant Design of Buildings \citep{BHRC2014}. Such improvements are reflected, for instance, in lower overall expected damage ratios, as reported by \citet{Ghodrati2013} for the city of Tehran. There is, however, room for improvement. The most recent of these studies, \citet{Khodaverdian_2016_BSSA}, presents an estimation of spatially-variable seismic parameters for Iran using an approach similar to the one we use here, but does not necessarily focus on hazard. Other previous efforts have outdated catalogs, use fault-centered hazard analyses, or focus on particular locations or cities.}

\myrevision{In the present study, we focus our attention on northern Iran. This region houses about 41 percent (32 million) of the total population of the country, with most living in the cities of Tehran, Mashhad, Tabriz, Karaj, Qom, Urmia, Rasht, and Ardabil. This region has suffered multiple devastating earthquakes in the past \citep[e.g.,][]{Mehrain_1990_Tech, Chafory-Ashtiany_1999_DPM, Razzaghi_2012_Tech}, and has recently been the subject of various efforts to better characterize seismic hazard \citep[e.g.,][]{Abdollahzadeh2014a, Boostan2015}. Among other things, these studies have highlighted the prevailing differences between alternative approaches, the importance of quantifying uncertainty, and the need for considering all potential faults in the region---even if believed inactive---by adopting new methodologies and the latest available data.}

\myrevision{Most previous studies for northern Iran focus on particular locations based on specific categorization of seismogenic faults \citep[e.g.][]{Ghodrati2003,Vafaie2011,Abdi2013}, which are often subjectively interpreted and limited to data availability. By contrast, we use an approach based on the smoothed seismicity method introduced by \citet{Frankel1995}. We use regionally-based and smoothed seismic parameters derived from an up-to-date seismic catalog, use an appropriate attenuation relationship, and consider various levels of uncertainty.} The smoothed seismicity approach has been used with success in other regions \citep[e.g.][]{Cao1996, Akinci2004, Kalkan2009}, including Iran \citep{Khodaverdian_2016_BSSA}. \myrevision{In \citep{Khodaverdian_2016_BSSA}, the seismic parameters $b$ and $M_{\max}$ are calculated for sites equally distributed all over the Iran on a grid 1*1 degree. Earthquake catalog for each grid point is all events within 200 km buffer zone surrounding the grid point. Whereas in this study, we assume the seismic parameters are uniform for each tectonic seismic region.} Its major feature is the treatment of cataloged instrumental and historical earthquakes as individual point sources as opposed to events that are part of a predefined system of faults in a given tectonic setting. While we recognize that \myrevision{fault-based/geologic-based earthquake recurrence estimates} can be more realistic and accurate, by choosing the smoothed seismicity method we seek to \myrevision{acknowledge} the uncertainty associated with fault mapping, earthquake location and source mechanism definitions, and to cover areas that would otherwise be underestimated or ignored. In our analysis, we assume earthquake hazard in northern Iran to be dominated by the seismic zones of Azerbaijan, Alborz, and Kopeh-Dagh, but also consider the contribution of the areas of Zagros and the central-east seismic zones to the hazard calculation within our domain of interest (see Fig.~\ref{fig:iran}). In addition to this five seismic-zone model, we obtain results for the entire region of interest as if it could be considered as a single uniform seismic zone. We used both these models and the combination of them to interpret seismic hazard in northern Iran.

In the following sections we describe our region of interest in greater detail, review its associated tectonics and seismicity, and provide additional information about the adopted approach. Significant aspects of our work include the composition of the earthquake catalog, including catalog completeness and the conversion of earthquake magnitudes, the selection of an appropriate attenuation relationship for the region, and the estimation of the Gutenberg-Richter seismicity parameters. Our analysis includes seismic hazard curves and maps for peak ground acceleration (PGA) considering exceedance probabilities of 2 and 10 percent in 50 years for rock site conditions for events with moment magnitude \myrevision{$M_w \geq 4.5$}. We analyze our results in light of similar estimates available in the literature and summarize the most relevant conclusions.