### 2016-10-20 hazard.tex
done for now with hazard section

\section{Seismic Hazard Approach}

There have been multiple assessments of seismic hazard for Iran. A brief report on the most recent review done to define the hazard zoning map of Iran included in the Iranian seismic code \citep{BHRC2014} can be found in \citet{Moinfar_2012_WCEE}. Other previous efforts include, for instance, the work of \citet{Tavakoli1999}, who presented iso-acceleration contour lines and seismic hazard zonation for Iran based on known major fault-lines and area-source models; the results obtained by \citet{Ghodrati2003}, who presented a refined hazard assessment for the metropolitan area of Tehran using a logic tree approach to combine results from three attenuation relationships; and the recent study by \citet{Khodaverdian_2016_BSSA}, who used a smoothed seismicity approach to estimate spatially variable seismicity parameters for the whole country.

Once a seismic catalog has been defined, the second most important step is the definition of the seismicity parameters $a$ and $b$ used in the Gutenberg-Richter recurrence relation
%
\begin{equation}
	\log_{10} N = a - b M \, ,
\end{equation}
%
\noindent
where $N$ is the number of earthquakes with magnitude greater than or equal to $M$, $a$ accounts for the number of earthquakes that occur within a given region, and $b$ is a measure of the occurrence of earthquakes relative to magnitude \citep{Gutenberg1944}. These parameters can be assumed constant within a region or spatially variable depending on the approach used.

In this study we use the smoothed seismicity method introduced by \citet{Frankel1995}. This method has been used successfully in different regions around the world including Alaska, California, Hawaii, Italy, \myrevision{and} Turkey \myrevision{\citep[e.g.,][]{Cao1996, Klein2001, Akinci2004, Kalkan2009, Moschetti2014,Kagan1994}}. In it, the $b$-value is commonly assumed constant within a given seismic zone of ``homogeneous'' seismogenic characteristics, whereas $a$ is varied spatially according to the cataloged occurrence of earthquakes within and around the region. The spatial variation of $a$ is defined based on dividing the region of interest into regular-size grid cells. The parameter $a$ is determined based on the number of events $n_i$ within each cell $i$ of magnitude greater than a reference value $M_{\mathrm{ref}}$, and discretized in different magnitude ranges (or bins). The number of earthquakes within each cell is then smoothed spatially taking into account the values of neighboring cells using a Gaussian function. In essence, the smoothed seismicity approach addresses the aleatoric uncertainty of future earthquake locations by using a weighted contribution of the background regional seismicity, and thus eliminates the need for knowledge about specific fault locations.

In this study we discretize the region of interest into a regular grid with cells of size 0.1\textdegree{} $\times$ 0.1\textdegree{} (in longitude and latitude). The parameter $a$ for each cell is obtained based on the information in the seismic catalog described in the previous section. Following \citet{Frankel1995}, $n_i$ is smoothed using the normalized Gaussian function
%
\begin{equation}
	\tilde{n}_i = \frac
		{ \sum_{j} n_{j} \exp \left( \frac{ -\Delta_{ij}^{2} }{ c^2 } \right) }
		{ \sum_{j} \exp \left( \frac{ -\Delta_{ij}^{2} }{ c^2 } \right) }
	\, ,
	\label{eq:nnorm}
\end{equation}
%
where $\tilde{n}_i$ is the smoothed, normalized number of events in cell $i$, $\Delta_{ij}$ is the distance between the $i$-th and $j$-th cells, and $c$ is the correlation distance. The summations in Eq.~\ref{eq:nnorm} are calculated over the number of $j$ cells within a distance $3c$ from the $i$-th cell. It should be noted that in this process we convert the cumulative number of events to incremental values following \citet{Herrmann1977}. We also note that \myrevision{other authors have used alternative smoothing kernels \citep[e.g.,][]{Moschetti2015} and that a comparison of the smoothing length between different methods may require additional considerations about the function decay. Last, we highlight that} according to \citet{Frankel1995}, moderate earthquakes generally occur in areas where there have been significant numbers of events with magnitudes $M \geq 3$, thus our interest in having augmented the earthquake catalog for events above this threshold.

It is obvious from Eq.~\ref{eq:nnorm} that an important parameter in smoothing the seismicity is the correlation distance. Different values of $c$ have been used in past studies, ranging between 10 and 50 km. \citet{Frankel1995} and \citet{Boyd2008} assumed a correlation distance of 50~km, \citet{Foteva2006} used 10 and 15~km, \citet{Barani2007} used 25~km, and \citet{Khodaverdian_2016_BSSA} used 40~km, \myrevision{while \citet{Boyd2007} and \citet{Petersen2015} include spatial variability in the smoothing parameter}. According to \citet{Mirzaei1997}, determining an appropriate value of $c$ depends on the crustal structure of the region, the distribution of stations, and the number and quality of records. \myrevision{Other studies employ likelihood testing of smoothed seismicity rate models to identify optimal smoothing lengths for adaptive and fixed smoothed seismicity methods \citep[e.g.,][]{Werner2011, Moschetti2015}}. We use an intermediate value of $c = 30$~km. This selection is consistent with epicentral uncertainty for previous hazard zoning maps for Iran \citep[see][]{Zare2012}.

In turn, we consider the $b$-values to be constant within each seismic zone as delimited in Fig.~\ref{fig:iran}. Details about the calculation of the $b$-value for each zone are provided \myrevision{in the next section}. By contrast, in the recent work by \citet{Khodaverdian_2016_BSSA}, the authors computed spatially-variable $b$-values within large grid-cells of size 1\textdegree{} $\times$ 1\textdegree{} based on the average seismicity within a 200-km radius from each cell. While gridding $b$-values and considering the influence of neighboring large-size cells allows \myrevision{one} to draw information from the background seismicity, it is unavoidable to have the contributing areas cross over seismic zones of different characteristics, which may bias the computation of $b$-values. Admittedly, these are subjective decisions and any reasonable choice can be considered valid or within the margins of the associated epistemic uncertainty.

Last, the annual rate $\lambda$ of exceeding a reference ground motion $u_o$ is defined by
%
\begin{align}
	\lambda & \left( u > u_o \right) = \nonumber \\
		& \sum_{k} \sum_{l} 10^{ {}^{ \left[ \log \left( \frac{ N_{k} }{ T } \right) - b \left( M_l - M_{\mathrm{ref}} \right) \right] } }
		P \left( u > u_o | D_k , M_l \right)
		\, .
	\label{eq:exceed}
\end{align}
%
Here, $N_k$ is the total of $\tilde{n}_i$ values for cells within a certain bin distance, where the values of $\tilde{n}_i$ have been previously binned by their distance from the site of interest; $T$ is the time in years used to determine $N_k$; $k$ and $l$ are the indexes of the distance and magnitude bins, respectively; and $P ( u > u_o | D_k , M_l )$ is the probability of $u$ exceeding $u_o$ given an earthquake of magnitude $M_l$ at a distance $D_k$ \citep{Frankel1995}.

It is clear from Eq.~\ref{eq:exceed} that the hazard calculations depend also on the choice of an attenuation relationship to predict the level of ground motion at a site ($u$) for any given earthquake. In the following two sections we describe how we obtained the $b$-values for the different seismic zones influencing the hazard in northern Iran and our choice of a suitable attenuation relationship (ground motion prediction equation).