### 2016-10-21 attenuation.tex
updates to sections 6 and 7

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-08}
    \caption{Variation of horizontal peak ground acceleration (PGA) with distance. KG2004: \citet{Kalkan2004}, BA2008: \citet{Boore2008}, AB2010: \citet{Akkar2010}, SO2012: \citet{Soghrat2012}. The dashed line are $\pm$uncertainty of KG2004.}
    \label{fig:att}
\end{figure*}

\section{Attenuation Relationship}

The last piece in our hazard analysis process is the choice of an adequate attenuation relationship, or ground motion prediction equation (GMPE). This is a highly influential factor because it controls \myrevision{the intensity level at a particular site given an earthquake of a certain magnitude and source characteristics and at a particular distance from the location of interest.}

Recently, \citet{Zafarani2014} investigated the predictive capabilities of a set of nine local, regional, and next generation attenuation (NGA) GMPEs to determine their applicability for northern Iran. They evaluated GMPE predictions against data from 32 earthquakes of magnitudes $M_w$ ranging between 4.7 and 7.4. This included comparisons of PGA and response spectral accelerations (SA) computed from time-series recorded on 163 stations located at epicentral distances of up to 200 km. The evaluation was done using the likelihood (LH) and log-likelihood (LLH) methods of \citet{Scherbaum_2004_BSSA, Scherbaum_2009_BSSA}. The combined evaluation for PGA, and SA for seven periods (T) between 0.1 and 2.0 s, yielded that the best overall predictions were those of the GMPEs introduced by \citet{Ghasemi_2009_JS}, \citet{Abrahamson_2008_ES}, and \citet{Chiou2008}. For the specific case of PGA ($T = 0$ s), however, the best results were those obtained with the GMPEs introduced by \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008}. \myrevision{Other GMPEs with good prediction results for the region were those of \citet{Soghrat2012} and \citet{Akkar2010}}.

\myrevision{A common approach used in seismic hazard analysis given the availability of multiple GMPEs is to address the associated epistemic uncertainty through a combination of the various GMPEs. However, in comparisons not shown here for brevity, we found that some of these GMPEs yielded mean predictions that were very close to each other. In an analysis based on the LLH coefficients reported by \citet{Zafarani2014}, and using the approach of \citet{Scherbaum_2009_BSSA} we found, for instance, that a logic-tree partition for the GMPEs proposed by \citet{Kalkan2004}, \citet{Chiou2008}, and \citet{Boore2008}, leads to weights of 0.3376, 0.3354, and 0.3270, respectively. These results meant that neither of these GMPEs has a substantial advantage over the others. This, however, would go against the need for accounting for uncertainty. Recently, \citet{Atkinson2014} argued that in such cases where multiple GMPEs do not offer sufficient variability, their combination does not necessarily and properly addresses uncertainty and suggested that, instead, one considers a set of high and low alternatives based on a central backbone model. Fig.~\ref{fig:att} illustrates the variation of horizontal PGA components with distance obtained using four of the GMPEs mentioned before for two different earthquake magnitudes. In these and other similar comparisons we noticed that the GMPE proposed by \citet{Kalkan2004} led to centered predictions and we chose this particular GMPE as our central backbone. Based on it, we defined an upper and lower bound (plus and minus one standard deviation, $\sigma = $????) as the two alternative GMPEs in our logic tree analysis, such that---for the most part---the amplitudes predicted by the mean values of other GMPEs fell within the range defined by these bounds.}

The selected backbone GMPE for our analysis, as proposed by \citet{Kalkan2004}, is of the form
%
\begin{align}
	\ln \left( Y \right) =
		& \hspace{1ex} b_1 + b_2(M_w - 6) + b_3 \left( M_w - 6 \right)^{2} \nonumber \\
		& + b_5 \ln \left( r \right) + b_V \ln \left( V_S / V_A \right)
	\, ,
\end{align}
%
where $Y$ is a ground motion parameter (here representing PGA), $V_A$ is a reference velocity in m/s, \myrevision{and} $V_S$ is the shear wave velocity of the site of interest. In this equation, the distance $r$ is given by
%
\begin{equation}
	r= \sqrt{ r^2_{\mathit{cl}} + h^2 }
	\, ,
\end{equation}
%
where $r_{\mathit{cl}}$ is the horizontal distance to the site of interest and $h$ is a reference fictitious depth, both given in km. According to \citet{Kalkan2004}, in the case of $Y$ representing PGA, the coefficients $b_1$, $b_2$, $b_3$, $b_5$ and $b_V$ are
%
\begin{equation}
\begin{array}{lcrlcr}
	b_1 &=&  0.393   \,,&\hspace{2em}   b_2 &=& 0.576\,,   \\
	b_3 &=& -0.107   \,,&\hspace{2em}   b_5 &=& -0.899\,,  \\
	b_V &=& -0.200   \,;
	\nonumber
\end{array}
\end{equation}
%
and $V_A$, $V_S$ and $h$ are
%
\begin{equation}
\begin{array}{lcrl}
	V_A &=& 1,112 & \mathrm{m/s}	\\
	V_S &=&   700 & \mathrm{m/s}	\\
	h   &=&  6.91 & \mathrm{km}\,,
	\nonumber
\end{array}
\end{equation}
%
respectively. The standard deviation of the residuals $(\sigma_{\ln y})$ expressing the random variability of the ground motions is 0.612. The value of $V_S$ is assumed to represent average surface rock sites.

We should note here that even though \citet{Kalkan2004} derived the above attenuation relationship for distances $r$ up to 250 km, \citet{Zafarani2014} only used data up to 200 km in their evaluation of the various GMPEs available. Therefore, to be consistent with the latter, we set the models in the hazard analysis to compute ground motion intensities (i.e., PGAs) at distances no greater than 200 km.